Bird beak's elaboration and exploration

### 2021-10-27 Analysis/02.1-shapespace_allbirds_lvl_superorder_order.Rmd
Create 02.1-shapespace_allbirds_lvl_superorder_order.Rmd

```{r install_beer, eval = FALSE}
## Installing beer
devtools::install_github("TGuillerme/elaboration_exploration_bird_beaks/beer")
```

```{r load_beer, eval = TRUE, message = FALSE}
set.seed(42)
## Loading beer
library(mcmcmcglmmm)
library(MCMCglmm)
library(dispRity)

## Loading the parametrisation data
load(file = "../Data/Processed/form_shape_space_1tree_2levels.rda")

## Select the data
shapespace <- form_shape_space_1tree_2levels$shape_space
groups_list <- form_shape_space_1tree_2levels$groups_list
selected_dim <- form_shape_space_1tree_2levels$shape_dim
consensus_tree <- form_shape_space_1tree_2levels$consensus_tree

@@

```{r, get_model_parameters, eval = FALSE}
## Load the computed models
param_chains <- list()
for(i in 1:3) {
    load(paste0("../Data/Cluster/allbirds_8748t_8Dshapespace_31levels/param_chains/params_shapespace_8D_31levels_20k_chain", i, ".rda"))
    param_chains[[i]] <- params_shapespace_8D_31levels[[1]]
}
## Extract the parameters from the parameter chains
est_params <- extract.parameters(param_chains)
```

@@

```{r, set_the_mini_chains, eval = TRUE}
## Get the distribution of trees
load(file = "../Data/Processed/trees_list.rda")

## The thinning parameter
thin <- 200
## The number of samples per chains
samples <- 10

## Set up the mini-chains models
shapespace_8748t_8D_32level <- make.mini.chains(
                                data         = shapespace,
                                dimensions   = selected_dim,
                                tree         = trees_list[sample(1:1000, 20)],
                                trait.family = "gaussian",
                                randoms      = c("global", "level1", "level2"),
                                residuals    = "global",
                                priors       = est_params$priors,
                                verbose      = TRUE,
                                parameters   = list(
                                    nitt   = est_params$burnin + thin * samples,
                                    thin   = thin,
                                    burnin = est_params$burnin))
## Save the mini.chains
save(shapespace_8748t_8D_32level, file = "../Data/Cluster/allbirds_8748t_8Dshapespace_31levels/shapespace_8748t_8D_32level.mini.chains")
```

@@

```{r}
## Preparing the model files
model_list <- list()
class(model_list) <- c("beer", "mini.chains")
models_path <- "../Data/Cluster/allbirds_8748t_8Dshapespace_31levels/cluster_out/"
## Loading all models
for(file in list.files(models_path)) {
    load(paste0(models_path, file, collapse = "/"))
    model_list[[length(model_list)+1]] <- model[[1]]
}

## Combining the model files
allbirds_8748t_8Dshapespace_31levels <- combine.mini.chains(model_list)

## Saving the results
save(allbirds_8748t_8Dshapespace_31levels, file = "../Data/processed/allbirds_8748t_8Dshapespace_31levels.rda")
```

<!--

#### Number of runs optimisation (TO BE DEVELOPED?)

```{r, eval = FALSE}
## Function for running one mini chain on a random tree (and saving the results)
fun.mini.chains <- function() {
    run.mini.chains(mini_chains, replicates = 1, path = "../Data/Processed/MCMC/", file.name = "mini.chains")
}
```

To make sure we reached consistent results, we run the set of N mini-chains were N was determined using the following variance optimisation algorithm.

 * 1. run an initial $N = 20$ chains;
 * 2. estimate the $V_{1}$, the pooled variances for each parameters in the chain (V1 is a vector of length §§§);
 * 3. update N to N = N * 1.05 and run an additional 5% chains;
 * 4. estimate $V_{2}$, the pooled variances for each parameters in the chain;
 * 5. calculate $\Delta_V$, the absolute change between $V_{1}$ and $V_{2}$ as ($\sqrt{(\frac{V_{1}}/{V_{2}} - 1)^{2}}$).
 * 6. If $\Delta_V < 0.05$ (meaning an increase of 5% of the replicates resulted in an increase of the parameters of less than 5%), the desired number of mini-chains is reached. Else, set $V_{1} = V_{2}$ and repeat from step 3.

```{r, optimised_replicates, eval = FALSE}
## The summary function

@@

After reaching the desired $N$ chains, we pulled all the posterior estimates together back into a "standard" big MCMCglmm with a posterior distribution including the effective results and the phylogenetic uncertainty.

-->

## References

>>>>

```{r load_packages, eval = TRUE, message = FALSE}
set.seed(42)
## Loading beer
library(mcmcmcglmmm)
library(MCMCglmm)
library(dispRity)

## Loading the parametrisation data
load(file = "../Data/Processed/shapespace_allbirds_lvl_superorder_order.rda")

## Select the data
shapespace <- shapespace_allbirds_lvl_superorder_order[[1]]$space
selected_dim <- shapespace_allbirds_lvl_superorder_order[[1]]$dimensions
consensus_tree <- shapespace_allbirds_lvl_superorder_order[[1]]$consensus_tree
trees_list <- shapespace_allbirds_lvl_superorder_order[[1]]$trees_list
```

@@

```{r, get_model_parameters, eval = FALSE}
## Load the computed models
param_chains <- list()
for(i in 1:3) {
    load(paste0("../Data/Cluster/shapespace_allbirds_lvl_superorder_order/param_chains/shapespace_allbirds_lvl_superorder_order_param", i, ".rda"))
    param_chains[[i]] <- shapespace_allbirds_lvl_superorder_order_param[[1]]
}
## Extract the parameters from the parameter chains
est_params <- extract.parameters(param_chains)
```

@@

```{r, set_the_mini_chains, eval = TRUE}
## The thinning parameter
thin <- 200
## The number of samples per chains
samples <- 10

## Set up the mini-chains models
shapespace_allbirds_lvl_superorder_order <- make.mini.chains(
                                data         = shapespace,
                                dimensions   = selected_dim,
                                tree         = trees_list[sample(1:1000, 1000)],
                                trait.family = "gaussian",
                                randoms      = c("global", "level1", "level2"),
                                residuals    = "global",
                                priors       = est_params$priors,
                                verbose      = TRUE,
                                parameters   = list(
                                    nitt   = est_params$burnin + thin * samples,
                                    thin   = thin,
                                    burnin = est_params$burnin))
## Save the mini.chains
save(shapespace_allbirds_lvl_superorder_order, file = "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/shapespace_allbirds_lvl_superorder_order.mini.chains")
```

@@

```{r}
## Preparing the model files
model_list <- list()
class(model_list) <- c("beer", "mini.chains")
models_path <- "../Data/Cluster/shapespace_allbirds_lvl_superorder_order/cluster_out/"
## Loading all models
for(file in list.files(models_path)) {
    load(paste0(models_path, file, collapse = "/"))
    model_list[[length(model_list)+1]] <- model[[1]]
}

## Combining the model files
shapespace_allbirds_lvl_superorder_order <- combine.mini.chains(model_list)

## Checking the diagnosis

## Saving the results
save(shapespace_allbirds_lvl_superorder_order, file = "../Data/processed/shapespace_allbirds_lvl_superorder_order.rda")

@@
